---
title: "Beyond Sequence-based Spatially-Resolved Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{beyond_visium}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  comment = "#>",
  fig.retina = NULL
)
```



Starting from Version 1.2.0, `escheR` package supports additional two data structures as input, including [`SpatialExperiment`](https://bioconductor.org/packages/release/bioc/html/SpatialExperiment.html) and `data.frame` from `base` R. In addition, `escheR` supports in-situ visualization of image-based spatially resolved data, which will be the focus of future development. 

# Visualized Dimensionality Reduced Embedding with `SingleCellExperiment`

## `SpatialExperiment` inherits `SingleCellExperiment`
Following the same syntax, one can also visualize dimensionality reduced embeddings of a `SpatialExperiment` object by providing the argument `dimred` with a non-null value. Hence, the first 2 columns of the corresponding `reducedDim(spe)` assay will be used as the x-y coordinate of the plot, replacing `spatialCoords(spe)`.

```{r redDim_spe}
library(escheR)
library(STexampleData)
library(scater)
library(scran)

spe <- Visium_humanDLPFC() |> 
  logNormCounts()
top.gene <- getTopHVGs(spe, n=500)

set.seed(100) # See below.
spe <- runPCA(spe, subset_row = top.gene) 

make_escheR(
  spe,
  dimred = "PCA"
) |> 
  add_fill(var = "ground_truth") +
  theme_minimal()
```

# Image-based `SpatialExperiment` Object
To demonstrate the principle that `escheR` can be used to visualize image-based spatially-resolved data pending optimization, we include two image-based spatially resolved transcriptomics data generated via seqFish platform and Slide-seq V2 platform respectively. The two datasets have been previously curated in the [`STexampleData`](https://bioconductor.org/packages/release/data/experiment/vignettes/STexampleData/inst/doc/STexampleData_overview.html) package

## seqFISH
```{r im_seqFISH}
library(STexampleData)
library(escheR)
spe_seqFISH <- seqFISH_mouseEmbryo()

make_escheR(spe_seqFISH) |>
  add_fill(var = "embryo")
```

> NOTE: trimming down the `colData(spe)` before piping into make-escheR could reduce the computation time to make the plots, specifically when `colData(spe)` contains extremely large number of irrelavent features/columns.


## SlideSeqV2
```{r im_slideseq}
library(STexampleData)
library(escheR)
spe_slideseq <- SlideSeqV2_mouseHPC()

make_escheR(spe_slideseq) |>
  add_fill(var = "celltype")
```

# Beyond Bioconductor Eco-system
We aim to provide accessibility to all users regardless of their programming background and preferred single-cell analysis pipelines. Nevertheless , with limited resource, our sustaining efforts will prioritize towards the maintenance of the established functionality and the optimization for image-based spatially resolved data. We regret we are not be able to provide seamless interface to other R pipelines such as `Seurat` and `Giotto` in foreseeable future.

Instead, we provide a generic function that works with a `data.frame` object as input. For example, relevant features in `Suerat` can be easily exported as a `data.frame` object manually or via `tidyseurat`[https://github.com/stemangiola/tidyseurat]. The exported data frame can be pipe into `escheR`.

```{r seurat_to_dataframe, eval = FALSE}
library(escheR)
library(Seurat)
pbmc_small <- SeuratObject::pbmc_small
pbmc_2pc <- pbmc_small@reductions$pca@cell.embeddings[,1:2]
pbmc_meta <- pbmc_small@meta.data

#> Call generic function for make_escheR.data.frame
make_escheR(
  object = pbmc_meta,
  .x = pbmc_2pc[,1],
  .y = pbmc_2pc[,2]) |> 
  add_fill(var = "groups")
```


# Session information

```{r}
utils::sessionInfo()
```
